---
title: "01_Bartels_Alan_2022_markdown"
author: "Alan Bartels"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
Load quakes with data(quakes). Show what’s there with str() and summary(). 
```{r quakes}
data(quakes)
str(quakes)
summary(quakes)
```
What do you see? 
Quakes is a data frame object with 1000 rows and five columns describing 
earthquakes. Each column is a vector describing a specific attribute of
earthquakes. These are labelled: lat, long, depth, mag, and station. These
vectors are classed as numeric, numeric, integer, numeric, and integer
respectively.

# Question 2
Show the entirety of column `long`
```{r long}
quakes$long
```

# Question 3
Apply `unique()`. What unique stations are there? Use `length()` with `unique()` to determine how many stations there are.
```{r unique}
unique(quakes$stations)
length(unique(quakes$stations))
```

# Question 4
Using `range()`, what is the range of depths where quakes occur?
```{r range}
range(quakes$depth)
```
# Question 5
Where was the earthquake of largest magnitude found? 
```{r max}
max_mag_index <- which(quakes$mag == max(quakes$mag))
quakes[max_mag_index,1:2]
```
Meta 1. For some of you, this week was your first experience with R (or even programming at all) - how did it feel to connect concepts to a novel set of problems. For those who have experience with R, was this a cakewalk or challenging? How did it feel to shake off the cobwebs?

This homework came pretty easily. Simple functions, variable setting, it feels good to get these cobwebs off. Pathing issues when I moved files around proved my greatest challenge, but I think I've gotten the hang of it now. 

Meta 2. How much time did this take you, roughly?

One hour. 

Meta 3. Please give yourself a weak/sufficient/strong assessment on this assignment. Feel free to comment on why.

I think I did sufficiently on this assignment, if I am able to get the github submission working I might give myself a strong. 

