---
title: "02_Bartels_Alan_2022"
author: "Alan Bartels"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load libraries, include = FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
library(ggridges)

theme_set(theme_bw(base_size=12))

ice <- read_csv("http://biol607.github.io/homework/data/NH_seaice_extent_monthly_1978_2016.csv") %>%
  mutate(Month_Name = factor(Month_Name),
         Month_Name = fct_reorder(Month_Name, Month))
```
# Part 2
```{r boxplots}
ice_extent <- ggplot(data = ice, 
                     mapping = aes(group = Month_Name,
                                   x = Extent,
                                   y = Month_Name))
ice_extent + geom_boxplot(fill = "turquoise")
```

# Part 3

```{r ridgelines}
ice_extent + geom_density_ridges(fill = "turquoise")
```
What do you learn that is different than the boxplot?

With ridgelines, one gets a visual of the frequency distribution of the sea ice
extent for a given month. For example, February and August do not look dissimilar
to other months in their box plot graph. However, in the ridgeline graph, one
can see that the distribution of sea ice extent during February and August 
appear to be bimodal, something that is not apparent in the box plot graph.

# Part 4
Making Groups from the Continuous
```{r monthly groups}
ice_by_month <- ggplot(data = ice, 
                       mapping = aes(
                         group = Month_Name,
                         x = Year,
                         y = Extent
                       ))
ice_by_month <- ice_by_month + geom_line()

ice_by_month + facet_wrap(vars(cut_interval(Month, n = 4)))
```

```{r annual groups}
annual_ice_extent <- ggplot(data = ice, 
                            mapping = aes(
                              group = Year,
                              x = Month_Name,
                              y = Extent
                            ))
annual_ice_graph <- annual_ice_extent + geom_line()

annual_ice_graph + facet_wrap(vars(cut_interval(Month, n = 4)))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
